\TheTitle

### 2019-07-25 Paper1/introduction.tex
results continued, intro rearranged a bit

Numerical methods for approximating the moments $\bm{\hat u}$ can be divided into intrusive and non-intrusive techniques. A popular non-intrusive method is the stochastic-Collocation (SC) method, see e.g. \cite{xiu2005high,babuvska2007stochastic,loeven2008probabilistic}, which computes the moments with the help of a numerical quadrature rule: For a given set of $Q$ quadrature weights $w_k$ and quadrature points $\bm{\xi}_k$, the moments are approximated by
\begin{align*}
\bm{\hat u}_i = \langle \bm{u}\varphi_i \rangle \approx \sum_{k = 1}^{Q}w_k \bm{u}({t,\bm{x},\bm{\xi}_k})\varphi_i(\bm{\xi}_k)f_{\Xi}(\bm{\xi}_k).
\end{align*} 
Commonly, SC uses sparse grids as quadrature rule, since they posses a reduced number of collocation points for multi-dimensional problems. Compared to tensorized quadrature sets, which require $O(M^s)$ quadrature points, sparse grids use $O(M(\log_2(M)^{s-1}))$ quadrature points to integrate polynomials of total degree $M$ exactly. 
Since the solution at a fixed quadrature point can be computed by a standard deterministic solver, the SC method does not require a significant implementation effort. Furthermore, SC is embarrassingly parallel, since the required computations can be carried out in parallel on different cores. A downside of collocation methods are aliasing effects, which stem from the inexact approximation of integrals. Therefore, collocation methods typically require more runs of the deterministic solver than intrusive methods \cite{xiu2009fast,alekseev2011estimation} to reach a given accuracy.

@@

When studying hyperbolic equations, the moment approximations of various methods such as Stochastic Galerkin \cite{le2004uncertainty}, IPM \cite{kusch2018filtered} and stochastic-Collocation \cite{barth2013non,dwight2013adaptive} tend to show incorrect discontinuities in certain regions of the physical space. These non-physical structures dissolve when the number of basis functions is increased \cite{pettersson2009numerical,offner2017stability} or when artificial diffusion is added through the spatial numerical method \cite{offner2017stability} or filters \cite{kusch2018filtered}. Also, a multi-element approach which divides the uncertain domain into cells and uses piece-wise polynomial basis functions to represent the solution has proven to mitigate non-physical discontinuities \cite{durrwachter2018hyperbolicity}. These structures commonly arise on a small portion of the space-time domain. Therefore, intrusive methods seem to be an adequate choice since they are well suited for adaptive strategies. By locally increasing the polynomial order \cite{tryoen2012adaptive,kroker2012finite,giesselmann2017posteriori} or adding artificial viscosity \cite{kusch2018filtered} at certain spatial positions and time steps in which complex structures such as discontinuities occur, a given accuracy can be reached with significantly reduced numerical costs. In addition to that, the number of unknowns when using intrusive instead of non-intrusive methods is significantly decreased in high dimensional problems: The number of moments $N$ given by \eqref{eq:numberBasisFcts} is asymptotically smaller than the number of quadrature points, which increases with $O(M(\log_2(M)^{s-1}))$. Therefore, one aim should be to accelerate intrusive methods, since they can potentially outperform non-intrusive methods in complex and high-dimensional settings. \\

@@

The effectiveness of these acceleration ideas are tested by comparing results with stochastic-Collocation for the uncertain NACA test case as well as a bent shock tube problem. Furthermore, we present a semi-intrusive method, which facilitates the task of implementing general intrusive methods. The method only requires the numerical flux of the deterministic problem (as well as an entropy if IPM is used). We thereby provide the ability to recycle existing implementations of deterministic solvers.

>>>>

Numerical methods for approximating the moments $\bm{\hat u}$ can be divided into intrusive and non-intrusive techniques. A popular non-intrusive method is the stochastic-Collocation (SC) method, see e.g. \cite{xiu2005high,babuvska2007stochastic,loeven2008probabilistic}, which computes the moments with the help of a numerical quadrature rule: For a given set of $Q$ quadrature weights $w_k$ and quadrature points $\bm{\xi}_k$, the moments are approximated by
\begin{align*}
\bm{\hat u}_i = \langle \bm{u}\varphi_i \rangle \approx \sum_{k = 1}^{Q}w_k \bm{u}({t,\bm{x},\bm{\xi}_k})\varphi_i(\bm{\xi}_k)f_{\Xi}(\bm{\xi}_k).
\end{align*} 
Commonly, SC uses sparse grids as quadrature rule, since they posses a reduced number of collocation points for multi-dimensional problems. Compared to tensorized quadrature sets, which require $O(M^s)$ quadrature points, sparse grids use $O(M(\log_2(M)^{s-1}))$ quadrature points to integrate polynomials of total degree $M$ exactly. 
Since the solution at a fixed quadrature point can be computed by a standard deterministic solver, the SC method does not require a significant implementation effort. Furthermore, SC is embarrassingly parallel, since the required computations can be carried out in parallel on different cores.

@@

When studying hyperbolic equations, the moment approximations of various methods such as Stochastic Galerkin \cite{le2004uncertainty}, IPM \cite{kusch2018filtered} and stochastic-Collocation \cite{barth2013non,dwight2013adaptive} tend to show incorrect discontinuities in certain regions of the physical space. These non-physical structures dissolve when the number of basis functions is increased \cite{pettersson2009numerical,offner2017stability} or when artificial diffusion is added through the spatial numerical method \cite{offner2017stability} or filters \cite{kusch2018filtered}. Also, a multi-element approach which divides the uncertain domain into cells and uses piece-wise polynomial basis functions to represent the solution has proven to mitigate non-physical discontinuities \cite{durrwachter2018hyperbolicity}. These structures commonly arise on a small portion of the space-time domain. Therefore, intrusive methods seem to be an adequate choice since they are well suited for adaptive strategies. By locally increasing the polynomial order \cite{tryoen2012adaptive,kroker2012finite,giesselmann2017posteriori} or adding artificial viscosity \cite{kusch2018filtered} at certain spatial positions and time steps in which complex structures such as discontinuities occur, a given accuracy can be reached with significantly reduced numerical costs. In addition to that, the number of unknowns when using intrusive instead of non-intrusive methods is significantly decreased in high dimensional problems: The number of moments $N$ given by \eqref{eq:numberBasisFcts} is asymptotically smaller than the number of quadrature points, which increases with $O(M(\log_2(M)^{s-1}))$. An additional downside of collocation methods are aliasing effects, which stem from the inexact approximation of integrals. Consequently, collocation methods typically require a higher number of unknowns than intrusive methods to reach a given accuracy \cite{xiu2009fast,alekseev2011estimation}. Therefore, one aim should be to accelerate intrusive methods, since they can potentially outperform non-intrusive methods in complex and high-dimensional settings. \\

@@

The effectiveness of these acceleration ideas are tested by comparing results with stochastic-Collocation for the uncertain NACA test case as well as a bent shock tube problem. Furthermore, we present a semi-intrusive method, which facilitates the task of implementing general intrusive methods. The method only requires the numerical flux of the deterministic problem (as well as an entropy if IPM is used). We thereby provide the ability to recycle existing implementations of deterministic solvers. Our numerical studies show the following main results:
\begin{itemize}
\item The standard IPM method cannot compete with SC and SG. However when using the discussed acceleration techniques, IPM reaches a given accuracy with the shortest run time.
\item Comparing SG with IPM, one observes that for the same number of unknowns, SG yields more accurate expectation values, whereas IPM leads to better variance approximations.
\item By studying aliasing effects, we show that SC requires a higher number of unknowns than intrusive methods, even for one spatial dimension to reach the same accuracy level.
\end{itemize}

### 2019-07-25 Paper1/results.tex
results continued, intro rearranged a bit

\subsection{2D Euler equations}
We start by quantifying the effects of an uncertain angle of attack $\phi\sim U(0.75,1.75)$ for a NACA0012 profile computed with different methods. The stochastic Euler equations in two dimensions are given by
\begin{align*}
\partial_t
\begin{pmatrix}
\rho \\ \rho v_1 \\ \rho v_2 \\ \rho e
\end{pmatrix}
+\partial_{x_1}
\begin{pmatrix}
\rho v_1 \\ \rho v_1^2 +p \\ \rho v_1 v_2 \\  v_1 (\rho e+p)
\end{pmatrix}
+\partial_{x_2}
\begin{pmatrix}
\rho v_2 \\ \rho v_1 v_2 \\ \rho v_2^2+p \\ v_2 (\rho e+p)
\end{pmatrix}
=\bm{0}.
\end{align*}
These equations determine the time evolution of the conserved variables $(\rho,rho \bm v, rho e)$, i.e. density, momentum and energy. A closure for the pressure $p$ is given by
\begin{align*}
p = (\gamma-1)\rho\left(e-\frac12(v_1^2+v_2^2)\right).
\end{align*}
Since the the fluid of the following test cases is air, we choose the heat capacity ratio $\gamma$ to be $1.4$. The spatial mesh discretizes the flow domain around the airfoil. At the airfoil boundary $\Gamma_{0}$, we use the Euler slip condition $\bm v^T\bm n = 0$, where $\bm n$ denotes the surface normal. At a sufficiently large distance away from the airfoil, we assume a far field flow with a given Mach number $Ma = 0.8$, pressure $p = 101,325$ Pa and a temperature of $273.15$ K. Now the angle of attack $\phi$ is uniformly distributed in the interval of $[0.75,1.75]$ degrees. I.e. we choose $\phi(\xi) = 1.25 + 0.5\xi$ where $\xi\sim U(-1,1)$.
The aim is to quantify the effects arising from the one-dimensional uncertainty $\xi$ with different methods. The IPM methods makes use of the acceleration strategies proposed in this work. To be able to measure the quality of the obtained solutions, we compute a reference solution using stochastic-Collocation with $100$ Gauss-Lobatto quadrature points.

@@

>>>>

\subsection{2D Euler equations with a one dimensional uncertainty}
We start by quantifying the effects of an uncertain angle of attack $\phi\sim U(0.75,1.75)$ for a NACA0012 profile computed with different methods. The stochastic Euler equations in two dimensions are given by
\begin{align*}
\partial_t
\begin{pmatrix}
\rho \\ \rho v_1 \\ \rho v_2 \\ \rho e
\end{pmatrix}
+\partial_{x_1}
\begin{pmatrix}
\rho v_1 \\ \rho v_1^2 +p \\ \rho v_1 v_2 \\  v_1 (\rho e+p)
\end{pmatrix}
+\partial_{x_2}
\begin{pmatrix}
\rho v_2 \\ \rho v_1 v_2 \\ \rho v_2^2+p \\ v_2 (\rho e+p)
\end{pmatrix}
=\bm{0}.
\end{align*}
These equations determine the time evolution of the conserved variables $(\rho,rho \bm v, rho e)$, i.e. density, momentum and energy. A closure for the pressure $p$ is given by
\begin{align*}
p = (\gamma-1)\rho\left(e-\frac12(v_1^2+v_2^2)\right).
\end{align*}
Since the the fluid of the following test cases is air, we choose the heat capacity ratio $\gamma$ to be $1.4$. The spatial mesh discretizes the flow domain around the airfoil. At the airfoil boundary $\Gamma_{0}$, we use the Euler slip condition $\bm v^T\bm n = 0$, where $\bm n$ denotes the surface normal. At a sufficiently large distance away from the airfoil, we assume a far field flow with a given Mach number $Ma = 0.8$, pressure $p = 101,325$ Pa and a temperature of $273.15$ K. Now the angle of attack $\phi$ is uniformly distributed in the interval of $[0.75,1.75]$ degrees. I.e. we choose $\phi(\xi) = 1.25 + 0.5\xi$ where $\xi\sim U(-1,1)$. As commonly done, the initial condition is equal to the far field boundary values. Consequently, the wall condition at the airfoil is violated and will correct the flow solution. The computational domain is a circle with a diameter of $40$ meters. In the center, the NACA0012 airfoil with a length of one meter is located. The discretization is composed of a coarsely refined far field and a finely resolved region around the airfoil, since we are interested in the flow solution at the airfoil. Altogether, the mesh consists of 22361 triangular elements.

The aim is to quantify the effects arising from the one-dimensional uncertainty $\xi$ with different methods. The IPM methods makes use of the acceleration strategies proposed in this work. To be able to measure the quality of the obtained solutions, we compute a reference solution using stochastic-Collocation with $100$ Gauss-Lobatto quadrature points. In the following, we investigate the L$^2$ error of the variance and the expectation value. Hence, when denoting the reference solution by $\bm u$, we have
\begin{align*}
\Vert E[\bm u] - E[\mathcal{U}] \Vert := \sqrt{\sum_{j=0}^{N_x} \Delta x_j \left( E[\bm u] - E[\mathcal{U}] \right)^2}, \enskip \Vert Var[\bm u] - Var[\mathcal{U}] \Vert := \sqrt{\sum_{j=0}^{N_x} \Delta x_j \left( Var[\bm u] - Var[\mathcal{U}] \right)^2}.
\end{align*}
The error is computed inside a box of one meter height and 1.1 meters length around the airfoil to prevent that small fluctuations in the coarsely refined far field influence the error.

@@

Recall that the numerical flux \eqref{eq:numericalFluxIPM} uses a quadrature rule to approximate integrals. We start by investigating the effects of this quadrature has on the solution accuracy. For this, we run the IPM method with a moment order ranging from three to seven \comment{[use IPM instead of osIPM, add accuracy of SC$^3$]} using a Clenshaw-Curtis quadrature rule with level three (i.e. 9 quadrature points) and level four (i.e. 17 quadrature points). The results are given in Figure TODO. It can be seen that the error stagnates when the chosen quadrature is not sufficiently accurate. This behavior results from aliasing effects, which dominate the accuracy level. Note that SC can be interpreted as an intrusive method which uses as many moments as quadrature points to compute the numerical flux. Therefore, we can conclude that the error of SC is dominated by aliasing effects and will not reach the same accuracy level as an intrusive method with an sufficiently accurate quadrature rule. Especially for high dimensional problems
